---
title: "tidying"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
options(tibble.print_min = 5)
```

```{r}
pulse_df <- haven::read_sas("data/public_pulse_data.sas7bdat") %>% 
    janitor::clean_names()

pulse_df = 
    pivot_longer(
        pulse_df,
        bdi_score_bl:bdi_score_12m,
        names_to = "visit",
        values_to = "bdi",
        names_prefix = "bdi_score_"
    )
```
To make this code chunk shorter and more organized, use the pipe operator

```{r}
pulse_df = haven::read_sas("data/public_pulse_data.sas7bdat") %>% 
    janitor::clean_names() %>% 
    pivot_longer(  
        bdi_score_bl:bdi_score_12m,
        names_to = "visit",
        values_to = "bdi",
        names_prefix = "bdi_score_"
    )
```

You could also change the value `bl` to `00m` for more consistency across visits. 
To do this you can use `replace`. Additionally, you want to convert `visits` to 
a factor variable to use levels in it (00m, 1m, 6m, 12m) as needed. Altogether, the code below will import, tidy, and transform the 
`pulse` dataset into a usable format.

```{r}
pulse_df = haven::read_sas("data/public_pulse_data.sas7bdat") %>% 
    janitor::clean_names() %>% 
    pivot_longer(  
        bdi_score_bl:bdi_score_12m,
        names_to = "visit",
        values_to = "bdi",
        names_prefix = "bdi_score_"
    ) %>% mutate(
        visit = replace(visit, visit == "bl", "00m"),
        visit = factor(visit)
    ) %>% arrange(id, visit)
```

# Learning assessment

```{r}
litters_data = read_csv("data/FAS_litters.csv") %>% 
    janitor::clean_names() %>% 
    select(litter_number, ends_with("weight")) %>% 
    pivot_longer(
        gd0_weight:gd18_weight,
        names_to = "gd",
        values_to = "weight"
    ) %>% mutate(
        gd = recode(gd, "gd0_weight"  = 0, "gd18_weight" = 18)
    )
```

## `pivot_wider`

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)
```

To change this to a wide format

```{r}
analysis_result_wide =
    pivot_wider(analysis_result,
                names_from = "time",
                values_from = "mean"
    )
```

## Binding rows

```{r}
fellowship_ring = readxl::read_excel("data/LotR_Words.xlsx",
                                     range = "B3:D6") %>% 
    mutate(movie = "fellowship_ring")

two_towers = readxl::read_excel("data/LotR_Words.xlsx",
                                     range = "F3:H6") %>% 
    mutate(movie = "two_towers")

return_king = readxl::read_excel("data/LotR_Words.xlsx",
                                     range = "J3:L6") %>% 
    mutate(movie = "return_king")
```

To combine all 3 data frame together, you could use `bind_rows` since they have 
the same variables

```{r}
lotr_tidy = bind_rows(fellowship_ring, two_towers, return_king)
```

Notice that there's a column for `male` and another for `female`; both could be 
pivoted under a new variable called `gender` and their values under a variable 
called `words`. additionally, you could change the values under the variable `race`
to lower case. 

```{r}
lotr_tidy = bind_rows(fellowship_ring, two_towers, return_king) %>% 
    janitor::clean_names() %>% 
    pivot_longer(
        female:male,
        names_to = "gender",
        values_to = "words"
    ) %>% mutate(
        race = str_to_lower(race)) %>% 
    select(movie, everything())
```

## Joining Datasets

Data can be spread across multiple related tables, in which case it is necessary
to combine or join them prior to analysis. 

```{r}
pups = 
    read_csv("data/FAS_pups.csv") %>% 
    janitor::clean_names() %>% 
    mutate(sex = recode(sex,`1` = "male", `2` = "female"),
           sex = factor(sex))

litter =
    read_csv("data/FAS_litters.csv") %>% 
    janitor::clean_names() %>% 
    separate(group, into = c("dose", "day_of_x"), sep = 3) %>% 
    mutate(
        wt_gain = gd18_weight - gd0_weight,
        dose = str_to_lower(dose)
    )

fas_data = left_join(pups, litter)
```

To return the all rows from x: `pups` without a match in y" `litter`:

```{r}
anti_join(pups, litter)
```

#Learning Assessment


